---
title: "Statistical Inference Assignment Part 2"
author: "agcode"
date: "24 December 2015"
output: html_document
---

---

## Tooth Growth Analysis 

---

### I. Overview

This exercise is going to analyze the ToothGrowth data in the R datasets package.

List of actions: 

- Load the ToothGrowth data and some basic exploratory data analyses 
- A basic summary of the data.
- Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.
- Conclusions and the assumptions needed for conclusions.

---

### II. Loading the ToothGrowth data and performing basic exploratory data analyses 

The following code loads the ToothGrowth dataset in R. After loading the dataset, some basic exploratory data analyses is performed:

```{r loadAndexplorechunk, echo=TRUE}

# Loading the dataset
data(ToothGrowth)
# Basic information about the dataset
str(ToothGrowth)

library(graphics)
  coplot(len ~ dose | supp, ToothGrowth, panel = panel.smooth, 
         show.given = FALSE, ylab = "Length", 
         xlab = c("Dose (mg)", 
                  "Tooth Growth given Orange Juice (Left) or Ascorbic Acid (Right)"))
```

---

### III. Summary of the data 

This dataset represents the Effect of Vitamin C on Tooth Growth in Guinea Pigs, according to the R Documentation. It contains the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

A data frame with 60 observations on 3 variables.

- [,1]   len	 numeric	 Tooth length
- [,2]	 supp	 factor	 Supplement type (VC or OJ).
- [,3]	 dose	 numeric	 Dose in milligrams.

```{r Summarychunk, echo=TRUE}
# Basic summary of the data
summary(ToothGrowth)
# Boxplot by supplement type and dose
boxplot(len ~ supp * dose, ToothGrowth, col="green", ylab="Tooth Length", xlab="Supplement & Dose",main="Boxplot by supplement type and dose") 
# Boxplot by supplement
boxplot(len ~ supp, ToothGrowth, col="blue", ylab="Tooth Length", xlab="Supplement type",main="Boxplot by supplement type") 

```

---

### IV. Compare tooth growth by supp and dose
Using confidence intervals and/or hypothesis tests, we compare tooth growth by supp and dose.

```{r confidencechunk, echo=TRUE}
# The T-test at 0.5 mg
t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == .5, ])
# The T-test at 1 mg
t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == 1, ])
# The T-test at 2 mg
t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == 2, ])

```

---

### V. Conclusions and Observations : 
 
The exploratory data analysis indicates that higher vitamin C dosages (in either orange juice[OJ] or pure ascorbic acid form[VC]) is an effective promoter of tooth growth.

The p-values generated by the T-tests above indicate that, for dosages of 0.5 mg and 1 mg, orange juice is more effective at promoting tooth growth.

Assumptions made: The observations are independent and identically distributed, random, the variances are different for the groups being compared in the t-tests, and the observations are unpaired.

---